### 2025-03-13 analysis/correction_example.qmd
correction: fixing intercept error

---
title: "correction_example"
format: pdf
---

```{r echo = FALSE, warning=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message = FALSE)
```

```{r}

library(tidyverse)
library(here)
library(lmerTest)
library(broom.mixed)
library(conflicted)
library(patchwork)

conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("lmer", "lmerTest")

```

```{r load-data}

miq_combined_filtered <- read.csv(here("data", "Combined", "Combined_filtered.csv"))
languages <- read.csv(here("data", "language.csv"))
color_scheme <- c("#85C0F9", "#A95AA1", "#F5793A")

```

```{r original-error, fig.width = 21.2, fig.height = 15}
#| fig.cap = "Original figure 4."

map_miq <- partial(map, .x = c("mdt", "mpt", "cabat") %>% set_names)

mods_old <- list()

extract_language_estimates <- function(mod_x) {
  fixed <- mod_x %>% tidy() %>% 
    filter(effect == "fixed",
           term %in% c("(Intercept)", "TonePitch-accented", "ToneTonal")) %>% 
    select(term, fixed_effect = estimate) %>%
    mutate(Tone = c("Non-tonal", "Pitch-accented", "Tonal"))

  # this adjusts the fixed-effect estimates so that non-tonal is a reference of 0
  fixed <- fixed %>%
    mutate(adj_fixed = fixed_effect - fixed %>% filter(Tone == "Non-tonal") %>% pull(fixed_effect))

  random <- tidy(mod_x, effects = "ran_vals", conf.int = TRUE, conf.level = 0.95) %>%
    filter(group == "language") %>%
    left_join(., languages %>% select(language, Tone), by = c("level" = "language")) %>%
    mutate(Tone = ifelse(is.na(Tone), "Tonal", Tone))

  combined <- left_join(random, fixed %>% select(-term), by = "Tone") %>%
    mutate(
      across(c(estimate, conf.low, conf.high), ~ .x + adj_fixed),
      language = fct_reorder(level, estimate)
    )
  return(combined)
}

run_mod <- function(term) {
  x <- reformulate("Tone*musicLessons + age + gender + (1|country) + (1|language)", response = term)
  mod <- lmer(x, miq_combined_filtered)
  output <- extract_language_estimates(mod)
  return(output)
}

mods_old$languages <- map_miq(~ run_mod(.x))

mods_old$languages$stats$mdt <- mods_old$languages$mdt |> 
  filter(Tone == "Tonal") |> 
  mutate(sig = ifelse(conf.low > 0, 1, 0)) |> 
  summarise(
    sig = sum(sig),
    n = n()
  )

mods_old$languages$stats$cabat <- mods_old$languages$cabat |> 
  filter(Tone == "Tonal") |> 
  mutate(sig = ifelse(conf.high < 0, 1, 0)) |> 
  summarise(
    sig = sum(sig),
    n = n()
  )

lang_counts <- miq_combined_filtered |> count(language)

lang_plotter2 <- function(data, x_lab, y_lab = NULL) {
  ggplot(data, aes(x = estimate, y = language, xmin = conf.low, xmax = conf.high, colour = Tone)) +
    geom_vline(xintercept = 0, lty = "dashed", colour = "grey50") + 
    geom_point(size = 2) +
    geom_errorbar(size = 1) +
    scale_colour_manual(values = c("#85C0F9", "#A95AA1", "#F5793A"),
                        labels = c("Non-tonal","Pitch-accented", "Tonal"),
                        name = "Language type") +
    lims(x = c(-0.9,0.9)) +
    labs(
      x = NULL,
      y = NULL, 
      title = x_lab
    ) +
    geom_text(aes(label = lang_label, x = conf.high + 0.1), colour = "black", hjust = 0, size = 6) +
    theme_minimal() +
    guides(colour = guide_legend(override.aes = list(size=10))) +
    coord_cartesian(clip = "off") +
    theme(axis.text.y= element_blank(),
          axis.title = element_text(size = 30),
          axis.ticks.y = element_blank(),
          axis.text.x = element_text(size = 16),
          legend.text = element_text(size = 22),
          legend.title = element_text(size = 24, hjust = 0.5),
          panel.grid = element_blank(),
          plot.title = element_text(hjust = 0.5, size = 24),
    ) 
}

format_language_label <- function(data, add_count = FALSE) {
  data <- data |> 
    mutate(level_clean = case_when(
      level == "Chinese/Mandarin" ~ "Mandarin",
      level == "Chinese/Cantonese/Yue" ~ "Cantonese",
      level == "Chinese/Other languages" ~ "Other Chinese languages",
      level == "Chinese/Taiwanese/Min Nan/Hokkien" ~ "Hokkien",
      TRUE ~ level
    ))

  if (add_count) {
    output <- data |>
    left_join(lang_counts, by = c("level" = "language")) |>
    mutate(n = as.integer(n)) |>
    mutate(lang_label = paste0(level_clean, " (", str_trim(format(n, big.mark = ",")), ")"))
  } else {
    output <- data |> 
      mutate(lang_label = level_clean)
  }
  return(output)
}

p1 <- lang_plotter2(format_language_label(mods_old$languages$mdt, TRUE), "Melodic Discrimination")
p2 <- lang_plotter2(format_language_label(mods_old$languages$mpt, TRUE), "Mistuning Perception")
p3 <- lang_plotter2(format_language_label(mods_old$languages$cabat, TRUE), "Beat Alignment")

p1+p2+p3 +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

ggsave(here("viz", "correction_figures", "original_fig4.png"), width = 20, height = 14)

```

```{r corrected-version, fig.width = 21.2, fig.height = 15}
#| fig.cap = "Fixing the intercept error."

map_miq <- partial(map, .x = c("mdt", "mpt", "cabat") %>% set_names)

mods_new <- list()

extract_language_estimates <- function(mod_x) {
  fixed <- mod_x %>% tidy() %>% 
    filter(effect == "fixed",
           term %in% c("ToneNon-tonal", "TonePitch-accented", "ToneTonal")) %>% 
    select(term, fixed_effect = estimate) %>%
    mutate(Tone = c("Non-tonal", "Pitch-accented", "Tonal"))

  # this adjusts the fixed-effect estimates so that non-tonal is a reference of 0
  fixed <- fixed %>%
    mutate(adj_fixed = fixed_effect - fixed %>% filter(Tone == "Non-tonal") %>% pull(fixed_effect))

  random <- tidy(mod_x, effects = "ran_vals", conf.int = TRUE, conf.level = 0.95) %>%
    filter(group == "language") %>%
    left_join(., languages %>% select(language, Tone), by = c("level" = "language")) %>%
    mutate(Tone = ifelse(is.na(Tone), "Tonal", Tone))

  combined <- left_join(random, fixed %>% select(-term), by = "Tone") %>%
    mutate(
      across(c(estimate, conf.low, conf.high), ~ .x + adj_fixed),
      language = fct_reorder(level, estimate)
    )
  return(combined)
}

run_mod <- function(term) {
  x <- reformulate("0 + Tone*musicLessons + age + gender + (1|country) + (1|language)", response = term)
  mod <- lmer(x, miq_combined_filtered)
  output <- extract_language_estimates(mod)
  return(output)
}

mods_new$languages <- map_miq(~ run_mod(.x))

mods_new$languages$stats$mdt <- mods_new$languages$mdt |> 
  filter(Tone == "Tonal") |> 
  mutate(sig = ifelse(conf.low > 0, 1, 0)) |> 
  summarise(
    sig = sum(sig),
    n = n()
  )

mods_new$languages$stats$cabat <- mods_new$languages$cabat |> 
  filter(Tone == "Tonal") |> 
  mutate(sig = ifelse(conf.high < 0, 1, 0)) |> 
  summarise(
    sig = sum(sig),
    n = n()
  )

p1 <- lang_plotter2(format_language_label(mods_new$languages$mdt, TRUE), "Melodic Discrimination")
p2 <- lang_plotter2(format_language_label(mods_new$languages$mpt, TRUE), "Mistuning Perception")
p3 <- lang_plotter2(format_language_label(mods_new$languages$cabat, TRUE), "Beat Alignment")

p1+p2+p3 +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

ggsave(here("viz", "correction_figures", "corrected_fig4.png"), width = 20, height = 14)

```